Tracing the Meta-Level: PyPy's Tracing JIT Compiler

### 2009-04-08 talk/icooolps2009/paper.tex
a final tweak

We attempt to use the technique of Tracing JIT Compilers
in the context
of the PyPy project, \ie on programs that are interpreters for some
dynamic languages, including Python.  Tracing JIT compilers can greatly
speed up programs that spend most of their time in loops in which they
take similar code paths.  However, applying an unmodified tracing JIT to
a program that is itself a bytecode interpreter results in very limited
or no speedup.
In this paper we show how to guide tracing JIT compilers to greatly
improve the speed of bytecode interpreters.  One crucial point is to
unroll the bytecode dispatch loop, based on two hints provided by the
implementer of the bytecode interpreter.  We evaluate our technique by
applying it to two PyPy interpreters: one is a small example, and the
other one is the full Python interpreter.

>>>>

We attempt to apply the technique of Tracing JIT Compilers
in the context
of the PyPy project, \ie to programs that are interpreters for some
dynamic languages, including Python.  Tracing JIT compilers can greatly
speed up programs that spend most of their time in loops in which they
take similar code paths.  However, applying an unmodified tracing JIT to
a program that is itself a bytecode interpreter results in very limited
or no speedup.
In this paper we show how to guide tracing JIT compilers to greatly
improve the speed of bytecode interpreters.  One crucial point is to
unroll the bytecode dispatch loop, based on two hints provided by the
implementer of the bytecode interpreter.  We evaluate our technique by
applying it to two PyPy interpreters: one is a small example, and the
other one is the full Python interpreter.
